---
title: "Final_Take_Home"
author: "Sumeet Mishra"
date: "4/28/2019"
output:
  html_document: default
  pdf_document: default
---

1.

Null Hypothesis, $H_0$ : Number of misprints follow Poisson distribution.
Alternate Hypothesis, $H_1$ : Number of misprints does not follow Poisson distribution.

```{r}
observed=c(9,23,40,30,31,26,19,10,5,4,3)
#plot(0:10, observed, type="h")
```

```{r}
num_obs=sum(observed)
total_misprints=sum((0:10)*observed)
ave=total_misprints/num_obs
expected=num_obs*dpois(0:20,ave)
data.frame(total_misprints=0:20,expected=round(expected,1))
```

Since the expected count for 8 misprints is less than 5. Hence, we combine expected count for 8 and above.


```{r}
#observed=c(9,23,40,30,31,26,19,10,5,7)
#expected=rep(NA,10)
#expected[1:9] = num_obs * dpois(0:8, ave) 
#expected[10] = num_obs * (1 - ppois(8, ave)) 
#data.frame(total_misprints=0:9,expected=round(expected,1))

observed=c(9,23,40,30,31,26,19,10,12)
expected=rep(NA,9)
expected[1:8] = num_obs * dpois(0:7, ave) 
expected[9] = num_obs * (1 - ppois(7, ave)) 
#data.frame(total_misprints=0:8,expected=round(expected,1))

G2 = 2 * sum(observed * log(observed/expected)) 
print(paste("Test Statistics of the likelihood test is",round(G2, 3)))
lh_p=1 - pchisq(G2, df=7)
print(paste("pvalue of the likelihood test is",round(lh_p, 3)))
```

Since pvalue from Likelihood test is greater than $\alpha$= 0.05, we fail to reject null hypothesis because we have insufficient evidence to conclude that the number of misprints does not follow Poisson distribution.

```{r}
X2 = sum((observed - expected)^2/expected) 
p_p=1 - pchisq(X2, df=7)

print(paste("Test Statistics of the Pearson test is",round(X2, 3)))

print(paste("pvalue of the Pearson test is",round(p_p, 3)))
```

Since pvalue from Pearson test is less than $\alpha$= 0.05, we reject null hypothesis because we have sufficient evidence to conclude that the number of misprints does not follow Poisson distribution.


2.

```{r}
examanxiety=read.table("examanxiety.txt",header=T)
#View(examanxiety)
```

a)

```{r}
x=examanxiety$Anxiety
y=examanxiety$Exam 
r = cor(x, y)
b = r * sd(y) / sd(x)
a = mean(y) - b * mean(x) 
#print(c(a,b))
yhat = a + b*((x))
print(paste("The regression line is given by, yhat =",round(a,3),"+", "(",round(b,3),")","* x"," which is in the form y_hat= intercept + slope * Anxiety" ))
```

```{r}
plot(x,y,main='Exam Score Vs Anxiety',xlab='Anxiety',ylab='Exam Score')
abline(a,b,col='red')
```


b)

Interpretation of Intercept:
When anxiety score is zero, then the predicted or mean exam score is 111.24.

Interpretation of Slope:
When anxiety score increases by one unit, the predicted or mean exam score decreases by the factor of slope i.e, 0.73.

Interpretation of yhat:
yhat is the regression line which represents the predicted exam score for a given anxiety score.

















