---
title: "Take Home Midterm 2"
author: "Sumeet Mishra"
date: "3/26/2019"
output:
  html_document: default
  pdf_document: default
---

```{r}
handwashing=read.table("handwashing.txt",header=T)
#View(handwashing)
```


##1.
## (a) What are the experimental units in this study?  What measurements are taken on each unit?  How many independent samples are there?

The experimental units are students from Michigan State. The measurements are the 6 moral judgement questions that each student is asked to give their responses to that is Dog, Trolley, Wallet, Plane, Resume and Kitten. There are 2 independent samples one is the handwashing sample group and the other is the control group(no handwashing group). 

## (b)  Perform a one-tailed significance test to study whether handwashing would lower the average answer to the trolley question.  Carefully define hypotheses, calculate a P-value, and write a substantive conclusion.


$H_0:\mu-\mu_1>=0$ \newline
$H_1:\mu-\mu_1<0$ \newline

where \newline
$\mu$= average answer of students from handwashing group \newline
$\mu_1$= average answer of students from control group

```{r}
hw=subset(handwashing,Condition==1)
#View(hw)
nhw=subset(handwashing,Condition==0)
#View(nhw)
x_hw=mean(hw$Trolley)
x_nhw=mean(nhw$Trolley)
s_hw=sd(hw$Trolley)
s_nhw=sd(nhw$Trolley)
n_hw=length(hw$Trolley)
n_nhw=length(nhw$Trolley)

test_stat=(x_hw-x_nhw)/sqrt((s_hw^2/n_hw)+(s_nhw^2/n_nhw))
df=((s_hw^2/n_hw)+(s_nhw^2/n_nhw))^2/(((s_hw^2/n_hw)^2/(n_hw-1))+((s_nhw^2/n_nhw)^2/(n_nhw-1)))
p_value=pt(test_stat,df)
p_value
```  

The p_value is calculated as  0.7181691 which is greater than $\alpha$ (0.05). Therefore we failed to reject null hypothesis and conclude that handwashing would not lower the average answer to the trolley question.

## (c)  Create a variable called Total that gives the total score on the six moral judgment questions for each individual.  (Remember to give R code.)  Draw labeled side-by-side  boxplots  (on  the  same  plot)  of Total for the handwashing and  control groups.

```{r}
handwashing$Total=(handwashing$Dog+handwashing$Trolley+
                     handwashing$Wallet+handwashing$Plane+handwashing$Resume+handwashing$Kitten)
total_hw=subset(handwashing,Condition==1)
#View(total_hw)
total_nhw=subset(handwashing,Condition==0)
#View(total_nhw)
boxplot(total_hw$Total,total_nhw$Total,
        main="Comparing Total counts for Handwashing and Control Groups", 
        ylab="Total",names=c('Handwashing Group','Control Group'))
```


## (d) Perform a one-tailed significance test to study whether handwashing would lower the average total score for the morality questions.Carefully define hypotheses,calculate a P-value, and write a substantive conclusion.

$H_0:\mu-\mu_1>=0$ \newline
$H_1:\mu-\mu_1<0$ \newline

where \newline
$\mu$= average answer of students from handwashing group \newline
$\mu_1$= average answer of students from control group

```{r}
x_total_hw=mean(total_hw$Total)
x_total_nhw=mean(total_nhw$Total)
s_total_hw=sd(total_hw$Total)
s_total_nhw=sd(total_nhw$Total)
n_total_hw=length(total_hw$Total)
n_total_nhw=length(total_nhw$Total)

test_total_stat=(x_total_hw-x_total_nhw)/sqrt((s_total_hw^2/n_total_hw)+
                                                (s_total_nhw^2/n_total_nhw))
df_total=((s_total_hw^2/n_total_hw)+(s_total_nhw^2/n_total_nhw))^2/
  (((s_total_hw^2/n_total_hw)^2/(n_total_hw-1))+((s_total_nhw^2/n_total_nhw)^2/(n_total_nhw-1)))
p_value=pt(test_total_stat,df_total)
p_value
```

The p_value is calculated as  0.3999633 which is greater than $\alpha$ (0.05). Therefore we failed to reject null hypothesis and conclude that handwashing would not lower the average total score for the morality questions.

## (e)  Find  95%  confidence  intervals  for  (i)  the  population  mean  of Total for  hand-washing, (ii) the population mean of Total for control, (iii) the difference in the population mean of Total between the two groups.

## (i) 95%  confidence  intervals  for the  population  mean  of Total for  hand-washing

```{r}
UL_hw=x_total_hw+qt(0.975,n_total_hw-1)*(s_total_hw/sqrt(n_total_hw))
LL_hw=x_total_hw-qt(0.975,n_total_hw-1)*(s_total_hw/sqrt(n_total_hw))
print(paste(" The 95% confidence interval for handwashing group for Total is between",
            LL_hw,"and",UL_hw))
```

## (ii) 95%  confidence  intervals  for the  population  mean  of Total for control

```{r}
UL_nhw=x_total_nhw+qt(0.975,n_total_nhw-1)*(s_total_nhw/sqrt(n_total_nhw))
LL_nhw=x_total_nhw-qt(0.975,n_total_nhw-1)*(s_total_nhw/sqrt(n_total_nhw))
print(paste(" The 95% confidence interval for control group for Total is between",
            LL_nhw,"and",UL_nhw))
```

## (iii) 95%  confidence  intervals for the difference in the population mean of Total between the two groups.

```{r}
UL_d_nhw=(x_total_hw-x_total_nhw)+qt(0.975,df_total)*sqrt((s_total_hw^2/n_total_hw)+
                                                            (s_total_nhw^2/n_total_nhw))
LL_d_nhw=(x_total_hw-x_total_nhw)-qt(0.975,df_total)*sqrt((s_total_hw^2/n_total_hw)+
                                                            (s_total_nhw^2/n_total_nhw))
print(paste(" The 95% confidence interval for the difference in the population mean of Total between the two groups is between",LL_d_nhw,"and",UL_d_nhw))
```

## (f)  Which  test  will  be  more  reliable  for  determining  whether  handwashing  affects moral judgments – the test for Trolley in part (b), or the test for Total in part (d)?  Explain.

The test we have done for total in part (d) will be more reliable. Because the total score takes count of all the scores across measurements (6 judgemental questions) in which Trolley is also included. In doing so, total experiment reduces any bias that was there for the trolley experiment.

##2. A  statistician  who  knows  much  more  about  cricket  and  rugby  than  basketball  and football  wishes  to  compare  the  size  of  NBA  and  NFL  players.   He  randomly  selects twelve players from each league and collects the data shown in the tables on the next page.

## (a) Find a 95% confidence intervals for: i.The mean height of all NBA players; ii.The mean height of all NFL players.

## (i) 95% confidence intervals for mean height of all NBA players

```{r}
height_NBA=c(80,82,82,77,74,81,83,80,76,82,81,81)
x_nba=mean(height_NBA)
s_nba=sd(height_NBA)
n_nba=length(height_NBA)
UL_nba=x_nba+qt(0.975,n_nba-1)*(s_nba/sqrt(n_nba))
LL_nba=x_nba-qt(0.975,n_nba-1)*(s_nba/sqrt(n_nba))
print(paste(" The 95% confidence interval for mean height of all NBA players is between",
            LL_nba,"and",UL_nba))
```

## (ii) 95% confidence intervals for mean height of all NFL players

```{r}
height_NFL=c(71,71,74,73,76,74,79,75,70,77,77,68)
x_nfl=mean(height_NFL)
s_nfl=sd(height_NFL)
n_nfl=length(height_NFL)
UL_nfl=x_nfl+qt(0.975,n_nfl-1)*(s_nfl/sqrt(n_nfl))
LL_nfl=x_nfl-qt(0.975,n_nfl-1)*(s_nfl/sqrt(n_nfl))
print(paste(" The 95% confidence interval for mean height of all NFL players is between",
            LL_nfl,"and",UL_nfl))
```

## (b)  (3 points) Find approximate 95% confidence intervals for: i.The median height of all NBA players; ii.The median height of all NFL players. If you cannot achieve exactly 95% confidence, get as close to that level of confidence as you can.

## (i) 95% confidence intervals for median height of all NBA players

```{r}
sort_height_NBA=sort(height_NBA)
alpha=1-0.95
k=qbinom(alpha/2,n_nba,0.5)
k
```
By trial and error k found to be 2 for 96% which is near to 95%  -
```{r}
1-2*pbinom(k,n_nba,0.5)
1-2*pbinom(k-1,n_nba,0.5)
kn=k-1
kn
print(paste("The 95% confidence interval for mean height of all NBA players is between",sort_height_NBA[kn+1],"and",sort_height_NBA[n_nba-kn]))

```

## (ii) 95% confidence intervals for median height of all NFL players

```{r}
sort_height_NFL=sort(height_NFL)
alpha=1-0.95
k=qbinom(alpha/2,n_nfl,0.5)
k
```

By trial and error k found to be 2 for 96% which is near to 95%  -
```{r}
1-2*pbinom(k,n_nfl,0.5)
1-2*pbinom(k-1,n_nfl,0.5)
kn=k-1
kn
print(paste("The 95% confidence interval for mean height of all NFL players is between",sort_height_NFL[kn+1],"and",sort_height_NFL[n_nba-kn]))

```


## (c) Suppose  we  are  willing  to  assume  that  both  NBA  heights  and  NFL heights have close to normal distributions.  Choose an appropriate statistical test for the null hypothesis that NBA players and NFL players have the same average height.  Calculate the test statistic and P-value, and explain in words what you may conclude from this analysis.

$H_0:\mu_0-\mu_1=0$\newline
$H_1:\mu-\mu_1\neq0$\newline

where $\mu_0$= average height of NBA players
      $\mu_1$= average height of NFL players



```{r}
test_stat_nn=(x_nba-x_nfl)/sqrt((s_nba^2/n_nba)+(s_nfl^2/n_nfl))
df_nn=((s_nba^2/n_nba)+(s_nfl^2/n_nfl))^2/
  (((s_nba^2/n_nba)^2/(n_nba-1))+((s_nfl^2/n_nfl)^2/(n_nfl-1)))
p_value=2*(1-pt(test_stat_nn,df_nn))
p_value
``` 


The p_value is calculated as  0.0000608201 which is less than $\alpha$ (0.05). Therefore we reject null hypothesis and conclude that the average height for NBA players and NFL players are not same.





