---
title: "StatsAss6"
author: "Sumeet Mishra"
date: "2/27/2019"
output: pdf_document
---

```{r}
x = scan("http://pages.iu.edu/~mtrosset/StatInfeR/Data/sample771.dat")
```
##Q1.(a) Empirical Plot of x

```{r}
plot(ecdf(x))
```

##1.(b)
##Mean

```{r}
mean(x)
```
##Var

```{r}
var=mean(x^2) - mean(x)^2
var
```

##1.(c)
##Median

```{r}
median(x)
```

##Interquantile range
```{r}
iq=quantile(x,0.75)-quantile(x,0.25)
iq
```

##1.(d)
##plug-in estimate of the interquartile range to the square root of the plug-in estimate of the variance

```{r}
iq/sqrt(var)
```

##1.(e)
##boxplot

```{r}
boxplot(x)
```

##1.(f)
##normal probability plot
```{r}
qqnorm(x)
```

##1.(g)
##kernel density estimate
```{r}
plot(density(x))
```

##1.(h)
##The sample was not drawn from normal distribution as it has two lumps and it's not symmetric(right skewed).

##Q.2.(a)

```{r}
am=read.csv('American_League.csv')
nat=read.csv('National_League.csv')
boxplot(am$W,nat$W,main = 'Boxplots of wins in 2015 for 2 leagues',
ylab='Number of wins',
names = c('American League', 'National League'))
```

##(b)
##Distribution of American league
```{r}
summary(am$W)
```

##Distribution of National league
```{r}
summary(nat$W)
```

```{r}
hist(am$W,main = 'Histogram of number of wins for American league',xlab='Number of wins')
```

```{r}
hist(nat$W,breaks=10,main = 'Histogram of number of wins for National league',xlab='Number of wins')
```


```{r}
plot(density(am$W), main='Kernel Density plot of American League')
```

```{r}
plot(density(nat$W), main='Kernel Density plot of National League')
```

##From the above graphs we can see the distributions of the wins of both American and National league. None of them are normal distributions but wins of National league is a bimodal distribution as it has 2 peaks in the density plots. Also from summary of both the distributions it is clear that National league has more wins than American league.



